---
title: "d13c precision"
output:
  html_document:
    df_print: paged
---

This notebook looks at historical d13C data for known-value samples. The intent is to look at the precision and accuracy of d13C over time.

## Initial goals

* Estimate the overall precision of d13C analysis at NOSAMS
* Investigate changes in precision over time
* Is there a difference between the Optima and Prism
* Is there a difference by sample type
* How does sample size affect precision and accuracy?


# Data

Initially, I'm using d13C data from the process tables. This gives rec_num, date, value, instrument. Need to add at least sample size. Where is the d13C split size recorded? 

Where's best for the datasets we use? I prefer to keep data off of github, so maybe depositing the data on the share drive and importing it to people's local folders? Alternatively, if people are set up to access the NOSAMS DB, we can use a script to update data.


load libraries

```{r, message=F, warning=F}
library(tidyverse)
library(skimr)
library(amstools)
```

Load the data and summarize.

```{r}
load("data/NOSAMS_d13c.rda")
skim(data)
```

Remove samples marked r&d

```{r}
data <- data %>%
  filter(rd == 0)
```

Find the sample types run the most.

```{r}
mostrun <- data %>%
  group_by(rec_num) %>%
  filter(n() > 1000)

mostrun %>%
  group_by(rec_num, name) %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())

ggplot(mostrun, aes(date, d13c)) +
  geom_point() +
  facet_wrap(vars(name))
```

There are a lot of outliers that will cloud what the variability for "normal" samples is.

## TIRI I

Look at TIRI-I.

```{r}
tirii <- data %>%
  filter(rec_num == 17185)
```

```{r}
tirii %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())

ggplot(tirii, aes(date, d13c)) +
  geom_point()
```

Remove obvious mislabels and problem samples...

```{r}
ggplot(tirii, aes(date, d13c, color = irms, shape = irms)) +
  geom_point() +
  ylim(-12, -8)
```

Plot by size of 13C split

```{r}
ggplot(tirii, aes(c13qty, d13c, color = irms, shape = irms)) +
  geom_point() +
  ylim(-12, -8) +
  xlim(0, 30)
```

Plot by total sample size (yield)

```{r}
ggplot(tirii, aes(tot_qty, d13c, color = irms, shape = irms)) +
  geom_point() +
  ylim(-12, -8) +
  xlim(0, 300)
```

Size of 13C split vs size of sample

```{r}
ggplot(tirii, aes(tot_qty, c13qty, color = irms, shape = irms)) +
  geom_point() +
  xlim(0, 300) +
  ylim(0, 30)
```


Remove outliers by interquartile distance and look at remaining data.

```{r}
tirii_no <- tirii %>%
  filter(!is.na(removeOutliers(d13c)))

tirii_no %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())

tirii_no %>%
  ggplot(aes(date, d13c, color = irms, shape = irms)) +
    geom_point() 
```

Is there a difference by instrument?

```{r}
tirii_no %>%
  group_by(irms) %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())

#t.test(d13c ~ irms, data = tirii)
anova(lm(d13c ~ irms, data = tirii))
```

## OX-I gas

Let's look at 149823, an OX-I gas.

```{r}
filter(data, rec_num == 149823) %>%
ggplot(aes(date, d13c)) +
  geom_point()
```

Remove obvious mislabels and problem samples...

```{r}
ox_no <- data %>%
  filter(rec_num == 149823,
         date > "2019-01-01",
         d13c < -15,
         d13c > -20) 

ox_no %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())

ox_no %>%
  ggplot(aes(date, d13c, color = irms, shape = irms)) +
    geom_point() 
```

0.06 permil. Not bad.

## All OX-I

Let's try and look at all the OX-I bulbs. Picking rec_nums with a consensus fm of 1.0398 (ox-I).

```{r}
oxstd <- standards %>%
  filter(fm_consensus == 1.0398) %>%
  select(rec_num)

ox <- data %>%
  filter(rec_num %in% oxstd$rec_num)

ox %>%
  ggplot(aes(date, d13c, color = as.factor(rec_num))) +
           geom_jitter()
```

And the summary for all things that are supposed to be OX-I

```{r}
ox %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())
```

Might be better to look at this by rec_num.

```{r}
ox %>%
  ggplot(aes(as.factor(rec_num), d13c, color = as.factor(irms))) +
           geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
  
ox %>%
  group_by(rec_num, irms) %>%
  summarize("d13C mean" = mean(d13c),
            "d13C sd" = sd(d13c),
            N = n())
```


