---
title: "NOSAMS δ^13^C"
output: html_document
---

TODO:

* Join with process tables to get δ^13^C qty, etc.

Some questions (with current answers):

* Do the data in mass_spec match those in the process tables?
* Does irms sample size affect precision or ratio? Not really. Still need to investigate ratio vs d13C split size and total sample size.
* How should we eliminate outliers to get at the precision and accuracy of d13C? Currently using 1.5 * IQR outside of 1st and 3rd quartiles.
* What samples should we use to estimate IRMS precision and accuracy? Using INST3 GAS and NBS-19.
* What samples should we use to estimate the same for the entire process? Using all samples in standards table.
* What can d13C tell us about fractionation in sample prep processes? There seems to be differences in SD between OC and HY samples. Need to add process codes.


## Load libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(amstools)
library(DBI)
library(skimr)
library(gt)
```

## Gather data

Create a database connection, and a dbplyr connection to mass_spec

```{r}
con <- conNOSAMS()
mass_spec <- tbl(con, "mass_spec")
```

Join to standards table to select standards. Some of these have a d13_cons value listed. Samples that are more than 1.5 * IQR above the first quartile or more than 1.5 * IQR below the third quartile are flagged as potential outliers. These statistics are calculated per rec_num.

```{r}
standards <- tbl(con, "standards")
d13c_data <- inner_join(mass_spec, standards) %>% 
  select(rec_num, ms_date, sample_num, sample_desc, sample_id, d13_cons, pres_rms, 
         "d13_meas" = c13_cr_1std, beam_size, mass_spec) %>%
  collect() 

# Add outlier flagging
d13c_data <- d13c_data %>% 
  group_by(rec_num) %>% 
  filter(!is.na(d13_meas)) %>% 
  mutate(outlier = is.na(removeOutliers(d13_meas))) %>% 
  ungroup() %>% 
  mutate(sample_type = ifelse(str_sub(sample_desc, 1, 2) %in% c("OC", "HY", "GS", "WS", "DO"), 
                              str_sub(sample_desc, 1, 2), 
                              NA))
```

## Table of standards

The table below has all standards measured more than 10 times. Outliers by IQR (see above) are not included.

```{r}
d13c_summary <- d13c_data %>% 
  group_by(sample_type, rec_num) %>%
  filter(n() > 10,
         outlier == FALSE) %>% 
  summarize(Name = str_trim(first(sample_id), side = "left"), #remove whitespace to list OX together
            "Consensus d13C" = first(d13_cons),
            "d13C_mean" = mean(d13_meas),
            "d13C_sd" = sd(d13_meas),
            N = n()) %>% 
  arrange(Name) 

d13c_summary %>% 
  gt() %>%
  fmt_number(
    columns = 5:6,
    decimals = 2,
    use_seps = FALSE
  ) %>% 
  tab_header(title = html("δ<sup>13</sup>C of NOSAMS and external standards"))

```


The average SD of IRMS samples grouped by rec_num is `r mean(d13c_summary$d13C_sd)`. This includes all samples with a rec_num in the standards table and with outliers removed groupwise by IQR. 

## IRMS standards

Below are data for INST3 GAS and NBS-19. INST3 GAS is an internal standard and has no inter-lab consensus value. The consensus value for NBS-19 is 1.95 ‰ (doi:10.1515/pac-2013-1023). This paper by T. Coplen et. al has a bunch of other consensus values as well.

### INST3 GAS

```{r}
inst3gas <- mass_spec %>% 
  filter(sample_desc == "INST3 GAS",
         !is.na(c13_cr_1std)) %>% 
  collect() %>% 
  mutate(outlier = is.na(removeOutliers(c13_cr_1std)))

inst3gas %>% 
  filter(outlier == FALSE) %>% 
  summarize("d13C mean" = mean(c13_cr_1std),
            "d13C sd" = sd(c13_cr_1std),
            N = n()) %>% 
  gt %>% 
  fmt_number(
    columns = 1:2,
    decimals = 2,
    use_seps = FALSE
  ) %>% 
  tab_header(title = "INST3 GAS")
```


### NBS-19

```{r}
nbs19 <- mass_spec %>% 
  filter(sample_desc %like% "NBS-19") %>%
  collect() %>% 
  mutate(outlier = is.na(removeOutliers(c13_cr_1std)))

nbs19 %>% 
  filter(outlier == FALSE) %>% 
  summarize("Consensus d13C" = 1.95,
            "d13C mean" = mean(c13_cr_1std),
            "d13C sd" = sd(c13_cr_1std),
            N = n()) %>% 
  gt %>% 
  fmt_number(
    columns = 2:3,
    decimals = 2,
    use_seps = FALSE
  ) %>% 
  tab_header(title = "NBS-19")
```

# Beam size vs precision/accuracy

Look at histogram of beam size

```{r}
mass_spec %>% 
  filter(beam_size > 0,
         beam_size < 0.001) %>%
  ggplot(aes(beam_size)) + geom_histogram()
```


INST3 Gas ratio vs beam size

```{r}
inst3gas %>% 
  filter(beam_size > 0,
         beam_size < 0.001,
         c13_cr_1std < -2) %>%
  ggplot(aes(beam_size, c13_cr_1std, color = outlier)) + 
           geom_point()
```

NBS-19 ratio vs beam size

```{r}
nbs19 %>% 
  filter(beam_size > 0,
         beam_size < 0.001,
         c13_cr_1std < 5,
         c13_cr_1std > -1) %>%
  ggplot(aes(beam_size, c13_cr_1std, color = outlier)) + 
           geom_point()
```

